---
title: "replication"
author: "DD"
date: "2021/10/27"
output: html_document
---


## Codebook 
Form: Variable name / Variable label / Variable values 



- agegroup / younger and older adult participants / factor, younger & older

- agegroup_num / younger and older adult participants / integer, 0 = younger, 1 = older

- covid_numbers / statement on regularly checking numbers of COVID infection, hospitalizations & deaths / 1 (Strongly agree) - 6 (strongly disagree)

- policies_UK / statement on how dissatisfied with UK government COVID policies / 1 (Extremely dissatisfied) - 6      (Extremely satisfied)

- rp_mean_norm / normalized scores on COVID risk perception / 0 (low risk perception) - 1 (high risk perception)

- dospert_beta / risk attitude scores / negative scores indicate risk averse, positive scores indicate risk seeking.

- objectiverisk_totalscore / Objective risk for COVID-19 serious illness or death / Higher scores indicating higher risk.

- numeracy_total / scores on the Lipkus Numeracy Scale / 0 (no correct answers) - 11 (all correct answers)

- numeracy3 / third item of numeracy scale, best predictor / 0 = incorrect, 1 = correct

- risktaking_mean / mean scores on COVID-19 risk-taking items / 0 (no risk-taking) - 5 (high risk-taking)

- risk-taking_enforced / subset of guidelines enforced by government / 0 (no risk-taking) - 5 (high risk-taking)

- risk-taking_recommendations / subset of guidelines unenforced by government / 0 (no risk-taking) - 5 (high risk-taking)

- rp_individual_mean / mean scores for personal risk perception items only / 0 (low risk perception) - 1 (high risk perception) 

- rp_individual_others / mean scores for perception of others' risk items only / 0 (low risk perception) - 1 (high risk perception) 







## 1. Planned analysis: mediation model

- 1.1 Step 1: H1 hypotheses
  Testing whether there is a significant relationship between x and y.
  Then to test whether the relationship between x and m is significant.

- 1.2 Step 2: H2 hypotheses
  Testing the relationship between predictor (x) and mediator (m)
  In this step, we assess whether there is a relationship between age and the chosen mediators. 
  
  
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages(ggplot2)
library(ggplot2)
#install.packages(tidyverse)
library(tidyverse)
#install.package(patchwork)
library(patchwork)
#install.packages(Hmisc)
library(Hmisc)
#install.packages("ggcorrplot")
library(ggcorrplot)
#install.package(psych)
library(psych)



coviddata <- read.csv("data_analysis.csv")
d <- read.csv("datastudy4.csv")



coviddata$agegroup_num <- coviddata$agegroup
coviddata$agegroup <- factor(coviddata$agegroup, levels = c("younger", "older"))
options(scipen=999)
options(digits=3)

```


## Descriptives
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
coviddata$agegroup_num[d$agegroup == "younger"] <- "0" 
coviddata$agegroup_num[d$agegroup == "older"] <- "1"

coviddata$agegroup_num <- as.integer(coviddata$agegroup_num)


correlationmatrix <- data.frame(coviddata$agegroup_num, coviddata$risktaking_mean,
                                coviddata$objectiverisk_totalscore,
                                coviddata$numeracy3, coviddata$dospert_beta)



corr <- cor(correlationmatrix, use = "complete.obs")

p.mat <- cor_pmat(correlationmatrix)



ggcorrplot(corr, hc.order = TRUE, type = "lower", lab = TRUE,
   outline.col = "white",
   ggtheme = ggplot2::theme_gray,
   colors = c("#E69F00", "white", "#0072B2"))

```


```{r}

descriptives <- data.frame(coviddata)

options(digits=4)

# all variables except dospert 
lapply(descriptives[, 3:14], mean)

lapply(descriptives[, 3:14], sd) 

# dospert mean and sd.
mean(descriptives$dospert_beta, na.rm = TRUE)
sd(descriptives$dospert_beta, na.rm = TRUE)

descriptives <- describeBy(coviddata, coviddata$agegroup) 
descriptives$younger
# just younger adults
descriptives$older# just older adults

# dospert mean and sd 
age_dospert <- coviddata %>%
  group_by(agegroup) %>%
  summarise(risktaking_mean = mean(dospert_beta, na.rm = TRUE),
            risktaking_sd = sd(dospert_beta, na.rm = TRUE))
age_dospert   

#frequencies for numeracy3
data.frame(table(coviddata$agegroup, coviddata$numeracy3))

```

## Overview of age group differences on COVID risk-taking items.
```{r}


data <- d %>% 
  dplyr::select(risktaking_handsanitizer, risktaking_touchface_rc, risktaking_handwashing,
                risktaking_crowdedspaces, risktaking_cleaning, risktaking_facemaskshops,
                risktaking_facemasktransport, risktaking_meetingindoors_rc, 
                risktaking_meterdistance, risktaking_deliveries, agegroup)

data <- data %>%
  rename("Hand sanitizer" = risktaking_handsanitizer,
         "Touching face" = risktaking_touchface_rc,
        "Washing hands" = risktaking_handwashing,
         "Avoiding crowded spaces" = risktaking_crowdedspaces,
        "Cleaning common surfaces" = risktaking_cleaning,
        "Wearing mask in shops" = risktaking_facemaskshops,
        "Wearing mask on transport" = risktaking_facemasktransport,
        "Meeting indoors" = risktaking_meetingindoors_rc,
        "Contact-free deliveries" = risktaking_deliveries,
        "Keeping meter distance" = risktaking_meterdistance)



data <- data %>% 
  pivot_longer(
    cols = c("Hand sanitizer","Touching face",
                      "Washing hands",
                      "Avoiding crowded spaces",
                      "Cleaning common surfaces",
                      "Wearing mask in shops",
                      "Wearing mask on transport",
                      "Meeting indoors",
              "Contact-free deliveries",
                      "Keeping meter distance"),
    names_to = "item", 
    values_to = "response",
    values_drop_na = TRUE
  )


cbPalette <- c("#E69F00", "#0072B2")


h1age <- ggplot(coviddata, aes(x = agegroup, y = risktaking_mean)) +
  geom_boxplot(aes(fill = agegroup)) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 3, fill = "black") +
  xlab("Age group") +
  ylab("COVID-19 risk-taking") +
   scale_fill_manual(values=cbPalette) +
   scale_y_continuous(limits = c(1, 5)) +
  theme(legend.position = "none")
print(h1age)

plot1 <- ggplot(data, aes(x = item, y = response)) +
  geom_boxplot(aes(fill = agegroup)) +
  xlab("Risk-taking items") +
  ylab("Response") +
  scale_y_continuous(limits = c(1, 5)) +
  scale_color_manual(values=cbPalette)+
  scale_fill_manual(values=cbPalette) +
  labs(fill = "Age group") +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        plot.margin = margin(0.5, 0, 0, 1, "cm"))
print(plot1)

ggsave("risktaking_items_oneplot.png", plot1, width = 10, height = 4)

ggsave("risktaking_items_oneplot.png", plot1 , width = 30, height = 15, dpi = 300, units = "cm")




```





## 1.1.1 Testing H1: age

In our registered report, we listed age as a H1 hypothesis, with the expectation that older adults are more likely to adhere to guidelines than younger adults. Age is the main variable in our paper and analysis. If age is not significantly related to COVID risk-taking, we will not run the remainder of the planned analysis, as this is a set requirement.

```{r H1: age, echo=TRUE}
cbPalette <- c("#E69F00", "#0072B2")

h1age <- ggplot(coviddata, aes(x = agegroup, y = risktaking_mean)) +
  geom_boxplot(aes(fill = agegroup)) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 3, fill = "black") +
  xlab("Age group") +
  ylab("COVID-19 risk-taking") +
   scale_fill_manual(values=cbPalette) +
   scale_y_continuous(limits = c(1, 5)) +
  theme(legend.position = "none")
print(h1age)

H1age <- lm(risktaking_mean ~ agegroup, data = coviddata)
summary(H1age)

```

Conclusion 1.1.1: Age group is a significant predictor of COVID risk-taking.This suggests that younger adults report adopting preventative behaviors less often. This also means that the planned analysis will proceed, as age was a set condition for the planned analysis.


## 1.1.2 Testing H1: objective risk

We included objective risk as we expect that those at higher risk of coronavirus complications will be more likely to
adhere to COVID 19 guidelines and implement health measures.


```{r H1: objective risk, echo=TRUE}

h1objectiverisk <- ggplot(coviddata, aes(x =objectiverisk_totalscore, y = risktaking_mean)) +  geom_jitter() +
  geom_smooth(method = "lm") +
    geom_point(alpha = 0.5) +
  xlab("Objective COVID-19 risk") +
  ylab("COVID-19 risk-taking") +
   scale_y_continuous(limits = c(1, 5)) 
print(h1objectiverisk)

H1objectiverisk <- lm(risktaking_mean ~ objectiverisk_totalscore, data = coviddata)
summary(H1objectiverisk)


```
Conclusion 1.1.2: Objective risk does predict protective covid behaviours.


## 1.1.3 Testing H1: risk attitude

We expect that those with an averse attitude towards health-related risk will be more likely to adhere to COVID-19 guidelines and implement health measures compared to those with a risk-seeking attitude.


```{r H1: risk attitude, echo=TRUE}

h1riskattitude <- ggplot(coviddata, aes(x =dospert_beta, y = risktaking_mean)) +  geom_jitter() +
  geom_smooth(method = "lm") +
    geom_point(alpha = 0.5) +
  xlab("Risk attitude") +
  ylab("COVID-19 risk-taking") +
   scale_y_continuous(limits = c(1, 5))
print(h1riskattitude)


H1riskattitude <- lm(risktaking_mean ~ dospert_beta, data = coviddata)
summary(H1riskattitude)

```

Conclusion 1.1.3: Self-reported health-related risk attitude does not predict COVID risky behavior.


## 1.1.4 Testing H1: numeracy

We expect that those with higher numerical ability will be more likely to adhere to COVID-19 guidelines and implement health measures compared to those with lower numerical ability.

```{r H1: numeracy, echo=TRUE}

cbPalette <- c("#E69F00", "#0072B2")

h1numeracy <- ggplot(coviddata, aes(x = as.factor(numeracy3), y = risktaking_mean)) +
  geom_boxplot(fill = cbPalette) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 3, fill = "black", position=position_dodge(width=0.75))+
  scale_x_discrete(limits = c("0", "1"),
                 labels = c("Incorrect", "Correct")) +
  xlab("Numeracy score") +
  ylab("COVID-19 risk-taking") +
  scale_y_continuous(limits = c(1, 5))
print(h1numeracy)


H1numeracy <- lm(risktaking_mean ~ numeracy3, data = coviddata)
summary(H1numeracy)

total_num_comparison1 <-lm(risktaking_mean ~ numeracy_total, data = coviddata)
  summary(total_num_comparison1)

```
Conclusion 1.1.4: Numeracy is a significant predictor of COVID risk-taking, suggesting that those with higher numerical abilities report adopting less preventative behaviors, thus taking more risk. 



## Combined plots of all H1 analyses.

```{r,message=FALSE,warning= FALSE}
h1age + h1objectiverisk + h1riskattitude + h1numeracy +
plot_layout(ncol = 4)

ggsave("combinedplots_h1.png", h1age + h1objectiverisk + h1riskattitude + h1numeracy +
plot_layout(ncol = 2), width = 6, height = 3.5)

ggsave("combinedplots_h1.png", h1age + h1objectiverisk + h1riskattitude + h1numeracy +
plot_layout(ncol = 2), width = 20, height = 15, dpi = 300, units = "cm")




```


# 1.2 Testing age group (x) and mediators (M)

In this section, we perform the second step of the mediation analysis, which is assessing whether there is a relationship between the predictor and the mediator(s). In our case, the predicting variable is age group, and the mediators are objective risk and numeracy. 


## 1.2.1 Objective risk and age
```{r}
obj_age <- lm(objectiverisk_totalscore ~ agegroup_num, data = coviddata)
summary(obj_age)

```
Conclusion: age and objective risk are related to one another; the older age group is more likely to have higher objective risk.


## 1.2.2 Numeracy and age 
```{r}

cbPalette <- c("#E69F00", "#0072B2")

age_num_h2 <- ggplot(coviddata, 
       aes(x = as.factor(numeracy3), y = ..prop.., 
           group = agegroup, fill = agegroup)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values=cbPalette) +
scale_x_discrete(limits = c("0", "1"),
                 labels = c("Incorrect", "Correct")) +
  scale_y_continuous(limits = c(0, 1)) +
  xlab("Numeracy score") +
  ylab("Proportion")
print(age_num_h2)

ggsave("age_numeracy_h2.png", age_num_h2, width = 10, height = 4)

AgeNumeracy <- lm(numeracy3 ~ agegroup_num, data = coviddata)
summary(AgeNumeracy)

# the analysis that includes the entire numeracy scale, as a check.
total_num_comparison2 <-lm(numeracy_total ~ agegroup_num, data = coviddata)
  summary(total_num_comparison2)

```
Conclusion: age and numeracy are related. The older age group has a lower numeracy score.



## 1.2.3 Mediation model 
The mediation model includes risk-taking, age group, and objective risk and numeracy as mediators
```{r}

# using bootstrapping approach to assess mediation effects.

mediation1 <- lm(risktaking_mean ~ agegroup_num + objectiverisk_totalscore + numeracy3, data = coviddata)
summary(mediation1)

mediation <- mediate(risktaking_mean ~ agegroup_num + (numeracy3) + (objectiverisk_totalscore), data = coviddata, n.iter=500)

print(mediation, short = FALSE)

```

## (The mediation analysis that includes the entire numeracy scale, as a check)
```{r echo=TRUE}

num_comparison <- mediate(risktaking_mean ~ agegroup_num + (numeracy_total) + (objectiverisk_totalscore), data = coviddata, n.iter=500)
print(num_comparison)
```






## 2. Exploratory analyses



## 2.1 Exploratory analyses: mediation model with age group (x), risk-taking (y) and risk perception (m).


```{r}

cbPalette <- c("#56B4E9", "#0072B2", "#D55E00", "#CC79A7")

general <- ggplot(coviddata, aes(x = agegroup, y = rp_mean_norm)) +
  geom_boxplot(aes(fill = agegroup)) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 3, fill = "black") +
   xlab("Age group") +
  ylab("Overall risk perception") +
  scale_fill_manual(values=cbPalette) +
   theme(legend.position = "none")
print(general)


# Step 1: checking the relationship between age group (x) and risk taking (y)

expl1 <- lm(risktaking_mean ~ agegroup, data = coviddata)
summary (expl1) 

# ...and risk perception (m) and risk taking.

expl1a <- lm(risktaking_mean ~ rp_mean_norm, data = coviddata)
summary (expl1a) 

# step 2: checking the relationship between age group (x) and risk perception (m)

expl2 <- lm(rp_mean_norm ~ agegroup, data = coviddata)
summary (expl2) 

# step 3: mediation

age_rp <- mediate(risktaking_mean ~ agegroup_num + (rp_mean_norm), data = coviddata, n.iter=500)

age_rp 




```




## 2.2 Risk perception for self, others and overall 

Instead of using all risk perception items, we have split them up by general, personal, and others. Perhaps there are differences in the risk perception for our own risk of COVID and the risk of others.

```{r}
# creating plots for perception of risk for self, and for others.
# the overall risk perception plot is already included in 2.2.

cbPalette <- c("#56B4E9", "#0072B2", "#D55E00", "#CC79A7")

individual <- ggplot(coviddata, aes(x = agegroup, y = rp_individual_mean)) +
  geom_boxplot(aes(fill = agegroup)) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 3, fill = "black") +
   xlab("Age group") +
  ylab("Risk perception for self") +
  scale_fill_manual(values=cbPalette) +
   theme(legend.position = "none")
print(individual)


others <- ggplot(coviddata, aes(x = agegroup, y = rp_other_mean)) +
  geom_boxplot(aes(fill = agegroup)) +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 3, fill = "black") +
   xlab("Age group") +
  ylab("Risk perception for others") +
  scale_fill_manual(values=cbPalette) +
   theme(legend.position = "none")
print(others)

general + individual + others

ggsave("combinedplots_age_rp.png", general + individual + others, width = 10, height = 4)

ggsave("combinedplots_age_rp_rt.png", general + individual + others, width = 20, height = 10, dpi = 300, units = "cm")


# Looking at risk-taking, risk perception, separated by age group.
# all risk perception items (individual, others, and general)

cbPalette <- c("#E69F00", "#0072B2")

plot4 <- ggplot(coviddata, aes(x = rp_mean_norm, y = risktaking_mean, 
                               shape = agegroup, colour = agegroup)) +
  geom_jitter() +
   geom_smooth(method = "lm") +
  xlab("Overall risk perception") +
  ylab("Risk-taking") +
   theme(legend.position = "none") +
   scale_colour_manual(values=cbPalette) +
   scale_y_continuous(limits = c(0, 5))
  


# only items refering to oneself
plot5 <- ggplot(coviddata, aes(x = rp_individual_mean, y = risktaking_mean, 
                               shape = agegroup, colour = agegroup)) +
  geom_jitter() +
   geom_smooth(method = "lm") +
  xlab("Risk perception for self") +
  ylab("Risk-taking")+
    theme(legend.position = "none") +
   scale_colour_manual(values=cbPalette) +
   scale_y_continuous(limits = c(0, 5))
  


# only items refering to others

legend_title = "Age group"

plot6 <- ggplot(coviddata, aes(x = rp_other_mean, y = risktaking_mean, 
shape = agegroup, colour = agegroup)) +
  geom_jitter() +
   geom_smooth(method = "lm") +
  xlab("Risk perception for others") +
  ylab("Risk-taking") +
   scale_colour_manual(legend_title, values=cbPalette) +
   scale_y_continuous(limits = c(0, 5)) +
  scale_shape_manual(name = legend_title, values = 16:17)
  



plot4 + plot5 + plot6






ggsave("combinedplots_age_rp_rt2.png", plot4 + plot5 + plot6 + plot_layout(ncol = 2) , width = 20, height = 15, dpi = 300, units = "cm")




```
The plots above show that there is a difference between how participants judge risk for ourselves and for others. The plots suggest that participants generally see more risk for others than themselves, even when they are at objectively higher risk for COVID serious consequences. When looking at age, younger adults seem to see more risk for themselves and others.


## 2.2.1 Mediation analysis on risk-taking, age group, and only risk perception for self.

```{r}

step1 <- lm(risktaking_mean ~ agegroup, data = coviddata)
summary (step1) 

step1a <- lm(risktaking_mean ~ rp_individual_mean, data = coviddata)
summary (step1a)

step2 <- lm(rp_individual_mean ~ agegroup, data = coviddata)
summary (step2) 

step3 <- lm(risktaking_mean ~ agegroup + rp_individual_mean, data = coviddata)
summary (step3) 

age_rp_individual <- mediate(risktaking_mean ~ agegroup_num + (rp_individual_mean), data = coviddata, n.iter=500)

age_rp_individual

```


## 2.2.2 Mediation analysis on risk-taking, age group, and only risk perception for others.
```{r}


step1 <- lm(risktaking_mean ~ agegroup, data = coviddata)
summary (step1) 

step1a <- lm(risktaking_mean ~ rp_other_mean, data = coviddata)
summary (step1a)

step2 <- lm(rp_other_mean ~ agegroup, data = coviddata)
summary (step2) 

step3 <- lm(risktaking_mean ~ agegroup + rp_other_mean, data = coviddata)
summary (step3) 

age_rp_other <- mediate(risktaking_mean ~ agegroup_num + (rp_other_mean), data = coviddata, n.iter=500)

age_rp_other

```



## 2.3 Age differences in COVID number checking and government COVID-19 approach dissatisfaction.

```{r, echo=TRUE}

# checking whether age groups differed in the descriptive items 
cbPalette <- c("#E69F00", "#0072B2")

# Checking COVID numbers

numbers <- ggplot(coviddata, aes(x= covid_numbers , fill = agegroup)) + geom_histogram(binwidth=1, binsize = 0.2, alpha=0.75,position="dodge")+
       xlab("Checking COVID-19 statistics") +
  ylab("Count") +
       scale_fill_manual(values=cbPalette) +
    theme(legend.position = "none")
print(numbers)

options(digits=8)

# younger adults report checking covid numbers significantly more than older adults.
t.test(covid_numbers ~ agegroup, data = coviddata)

# in this case, extremely satisfied is "1", extremely dissatisfied is "7".

policies <- ggplot(coviddata, aes(x= policies_UK , fill = agegroup)) + geom_histogram(binwidth=1, binsize = 0.2, alpha=0.75,position="dodge")+
         xlab("Dissatisfaction UK COVID-19 policies") +
  ylab("Count")+
       scale_fill_manual(values=cbPalette) +
  labs(fill = "Age group")
print(policies)

# younger adults report being significantly more dissatisfied with the government's approach to stop
# the spread of coronavirus.
t.test(policies_UK ~ agegroup, data = coviddata)

numbers+ policies

ggsave("numbers_policies.png", numbers+ policies, width = 10, height = 4)

ggsave("numbers_policies.png", numbers+ policies, width = 20, height = 10, dpi = 300, units = "cm")

```